---
title: "Lab1"
author: "RawanGalal"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load}
library(tidyverse)

```

```{r }
data = read_csv("LAB 1 - Online Retail Data.csv")
data
```

# 1. Data validation
## Level 1 –Cells, records, file
```{r}
summary(data)
```
# 2. DATA CLEANSING
## Handling negative values and NA's in UnitPrice:
### about 3.5% of the UnitPrice column is missing, and 1.5% are nulls.
### I decided to drop these values. If the percentage was higher, I would fill it with the average of the unit price, per stockcode.
```{r}
data %>%
  filter(UnitPrice > 0) -> data
data
###droped negative rows(11, 574)

```
## Remove duplicate rows
```{r}
data %>% distinct() %>% arrange(InvoiceNo) -> data
#unique(data)


```

```{r}
summary(data)
```

## IQR for UnitPrice is 3
```{r}
#data %>% 
 # filter(!is.na(UnitPrice)) -> data
  #hist(to_integer(data["UnitPrice"]))

### mutate(across(is.numeric, ~ replace_na(., 0) )) 
```


```{r}

count(data["UnitPrice"])


```

  
### null values successfully removed!


## Detecting outliers
```{r}
  hist( as.integer(data$UnitPrice), xlim=c(0,40000), breaks=20, ylim=c(0,50)) 

```
### As the histogram illustrates, there's 1 observation that is considered an outlier: the maximum UnitPrice = 38970.

```{r}
data %>% arrange(desc(UnitPrice))
```
### After further investigation in the data, it turns out that this extreme price corresponds to customer number 15098. This customer performed another transaction with a total price of 297. This is interesting! It shows an unusual behavior for the customer. It could indicate a fraud. So I decided to keep this extreme point.

### Now, we're ready to answer Q1.

```{r}
##data %>%
  ##group_by(StockCode)-> grouped_data

##grouped_data%>%
  ##summarise( avg_price = mean(UnitPrice),
    ##          UnitPrice= UnitPrice, 
      ##       UnitPrice_ = replace_na(UnitPrice, avg_price))-> data_



```


## Q1. What is the average price for items shipped international Vs Domestic?

```{r}
data %>%
  group_by(Destination) %>%
  summarize(avg_price = mean(UnitPrice * Quantity))

```

### clearly, items shipped internationally contribute more to our revenue. This might open doors for methods to sustain international customers, or increase the rate of purchase for domestic ones.

### ###############################################
## Q2: What is the most sold product category?

## Detecting outliers in the Quantity column
```{r}
  hist( as.integer(data$Quantity), xlim=c(0,80000), breaks=30, ylim=c(0,40)) 

```

### clearly, there's one extreme point (the Max. Quantity, which equals 80995.00). This point corresponds to a **paper craft, little birdie** product. Again, interesting. Could indicate a fraud, (For a theif throwing a party perhabs!) since that customer only has another transaction where he bought 1 pantry scrubbing brush. or perhaps the customer just started his crafts business! In either cases, This point is too important to exclude from the analysis!
```{r}
data %>%
  group_by(StockCode) %>%
  mutate(quantity_sold = if_else(Quantity>0, Quantity, 0)) %>%
  summarise(quantity_per_cat = sum(quantity_sold)) %>%
 arrange(desc(quantity_per_cat))


# sort / take the maximum
```


## -  Answer:  the most sold product category is *23843* which we sold 80995 items from it between 01/12/2010 and 09/12/2011 

# ##############################
## Q3: What is the most profitable product category?
### The Revenue column doesn't contain any null values.


```{r}
data %>%
  group_by(StockCode) %>%
  summarize(revenue_per_category = sum(Revenue)) %>%
  arrange(desc(revenue_per_category))
```

## - Answer: the most profitable product category is *23843*. This category alone increases our revenue by *168469.60*
## (Technically, Profit = revenue - cost. But since the information of cost isn't available, I used the revenue column only.)

# ####################################################
## Q4: 4. Which country has the highest cancellation rates .. is it really concerning?

## Detecting outliers for the canceled column.
```{r}
  hist( as.integer(data$Canceled), xlim=c(0,2), breaks=20, ylim=c(0,500)) 
```

### No outliers!

```{r}

##group by country 
data %>%
  group_by(Country) %>%
  summarize(total_cancellation = sum(Canceled), 
            quantity_sold = sum(Quantity)) %>%
  arrange(desc(total_cancellation))
```

## - Answer: the country which has the highest cancellation rates is *United Kingdom* , with total number of cancellations of 4408 + 179 +11 = 4598. 
## Although this might seem like alot, but this shouldn't be concerning, since the ration of the total_cancellation to the total items sold in the UK is less that .0002!
## On the other hand, US has a cancellation/sold ration of 0.2. This raises questions and could require further investigation.